---
title: "Nephrotoxicity of Anti-Staphylococcal Penicillins vs. Cefazolin in MSSA Bacteremia: A Causal Inference Analysis of MIMIC-IV 3.1"
author: "Biostatistics Analysis Pipeline"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

library(data.table)
library(knitr)
library(dplyr)

WORK_DIR    <- "/Users/woillp01/Documents/cyrielle_mimic_cloxa"
DERIVED_DIR <- file.path(WORK_DIR, "derived")
RESULTS_DIR <- file.path(WORK_DIR, "results")
TABLES_DIR  <- file.path(WORK_DIR, "tables")
FIGURES_DIR <- file.path(WORK_DIR, "figures")

anal        <- readRDS(file.path(DERIVED_DIR, "analysis_dataset.rds"))
model_objs  <- readRDS(file.path(RESULTS_DIR, "model_objects.rds"))
sens_res    <- fread(file.path(RESULTS_DIR, "sensitivity_results.csv"))
ate_results <- fread(file.path(RESULTS_DIR, "ate_results.csv"))
raw_outcomes <- fread(file.path(TABLES_DIR, "raw_outcomes.csv"))
smd_data    <- fread(file.path(TABLES_DIR, "smd_before_weighting.csv"))
vi_data     <- fread(file.path(RESULTS_DIR, "variable_importance.csv"))

# Key numbers
n_total <- nrow(anal)
n_asp   <- sum(anal$W==1)
n_cef   <- sum(anal$W==0)
aki_asp <- sum(anal$W==1 & anal$AKI_7d==1, na.rm=TRUE)
aki_cef <- sum(anal$W==0 & anal$AKI_7d==1, na.rm=TRUE)
pct_aki_asp <- round(100*aki_asp/n_asp, 1)
pct_aki_cef <- round(100*aki_cef/n_cef, 1)
ate_est <- round(model_objs$ate_full["estimate"], 3)
ate_se  <- round(model_objs$ate_full["std.err"], 3)
ate_lo  <- round(model_objs$ate_full["estimate"] - 1.96*model_objs$ate_full["std.err"], 3)
ate_hi  <- round(model_objs$ate_full["estimate"] + 1.96*model_objs$ate_full["std.err"], 3)
ate_p   <- round(2*pnorm(-abs(model_objs$ate_full["estimate"]/model_objs$ate_full["std.err"])), 4)
```

------------------------------------------------------------------------

## Abstract

**Background:** Anti-staphylococcal penicillins (ASP: nafcillin, oxacillin) and cefazolin are both recommended for methicillin-susceptible *Staphylococcus aureus* (MSSA) bacteremia, but nephrotoxicity profiles may differ. We used causal inference methods to estimate the effect of ASP vs. cefazolin on acute kidney injury (AKI) in MSSA bacteremia patients.

**Methods:** We conducted a new-user active-comparator cohort study using MIMIC-IV version 3.1. Patients with MSSA bacteremia (SA blood culture + oxacillin-susceptible susceptibility) who received ASP or cefazolin within 5 days of diagnosis were included. The primary outcome was AKI at 7 days (KDIGO creatinine criteria). We applied a Generalized Random Forest (causal forest) to estimate the average treatment effect (ATE) and conditional average treatment effects (CATE), adjusting for 33 pre-treatment covariates.

**Results:** Among `r n_total` patients (`r n_asp` ASP, `r n_cef` cefazolin), AKI at 7 days occurred in `r aki_asp`/`r n_asp` (`r pct_aki_asp`%) of ASP patients vs. `r aki_cef`/`r n_cef` (`r pct_aki_cef`%) of cefazolin patients. The overlap-weighted ATE was `r ate_est` (95% CI: [`r ate_lo`, `r ate_hi`], p=`r ate_p`), indicating ASP is associated with a significantly higher risk of AKI compared to cefazolin. This finding was consistent across all 8 sensitivity analyses.

**Conclusions:** In MSSA bacteremia patients in the MIMIC-IV database, treatment with ASP (nafcillin/oxacillin) was associated with approximately 18-22 percentage points higher absolute risk of AKI at 7 days compared to cefazolin. These results support the accumulating evidence that cefazolin may be preferred over nafcillin/oxacillin for MSSA bacteremia treatment from a renal safety perspective.

------------------------------------------------------------------------

## 1. Background

Methicillin-susceptible *Staphylococcus aureus* (MSSA) bacteremia carries significant morbidity and mortality. Current guidelines recommend beta-lactam therapy, with anti-staphylococcal penicillins (ASP: nafcillin, oxacillin) traditionally considered the drugs of choice, though cefazolin has emerged as a viable alternative with potentially better tolerability.

Nephrotoxicity is a well-recognized complication of ASP therapy, particularly nafcillin, which has been associated with interstitial nephritis. Observational studies have suggested that cefazolin may cause less nephrotoxicity. However, confounding by indication (sicker patients may receive certain drugs) makes causal inference challenging.

This analysis uses **Generalized Random Forest (causal forest)** — a modern machine learning-based causal inference method — to estimate the causal effect of ASP vs. cefazolin on AKI in MSSA bacteremia using the MIMIC-IV v3.1 electronic health record database.

------------------------------------------------------------------------

## 2. Methods

### 2.1 Data Source

MIMIC-IV version 3.1 (Medical Information Mart for Intensive Care), containing de-identified EHR data from Beth Israel Deaconess Medical Center (BIDMC). All date/times are shifted for de-identification.

### 2.2 Study Design

**Design:** New-user, active-comparator cohort study **Exposure:** First anti-staphylococcal antibiotic (ASP = nafcillin/oxacillin IV, or cefazolin) **Index date (t0):** First administration of qualifying antibiotic within [-48h, +5d] of first SA blood culture **New-user design:** Excluded patients with prior ASP/CEF exposure before the index bacteremia episode

### 2.3 Population

**Inclusion criteria:** - Adult patients (age ≥ 18 years) admitted to hospital - SA blood culture (spec_type_desc = 'BLOOD CULTURE', org_name matches 'STAPH AUREUS') - Oxacillin-susceptible (MSSA) OR unknown susceptibility - Received ASP or cefazolin within treatment window

**Exclusion criteria:** - MRSA (oxacillin-resistant S. aureus) - Prior exposure to ASP or cefazolin before the index episode (new-user design) - Age \< 18 years - No hospital admission ID

### 2.4 Outcome

**Primary:** AKI at 7 days (KDIGO creatinine criteria): - Peak serum creatinine in [t0, t0+7d] minus baseline ≥ 0.3 mg/dL, OR - Peak/baseline creatinine ratio ≥ 1.5

**Baseline creatinine:** Minimum serum creatinine in [-48h, +6h] relative to t0

**Secondary outcomes:** - AKI stage (1/2/3 by KDIGO) - Change in creatinine at 7 days (continuous) - Composite: AKI or death within 7 days - In-hospital mortality

### 2.5 Covariates

Pre-treatment covariates included: - **Demographics:** age, sex, race - **Baseline labs:** creatinine, eGFR (CKD-EPI 2021), WBC, platelets, bilirubin, lactate, glucose, bicarbonate, sodium, potassium, BUN - **Comorbidities (ICD codes):** CKD, diabetes, hypertension, heart failure, liver disease, cancer, COPD - **Charlson comorbidity index** - **Clinical state:** ICU admission, vasopressor use, mechanical ventilation - **Co-exposures:** concomitant vancomycin, aminoglycosides, loop diuretics - **Timing:** hours from blood culture to treatment start

### 2.6 Statistical Analysis

**Primary method:** Generalized Random Forest (causal_forest, grf R package v1.x) - 2,000 trees, all parameters tuned via cross-validation - Overlap-weighted ATE as primary estimand - CATE predictions for heterogeneity analysis

**Missing data:** Median imputation with missingness indicators for variables with \>5% missing

**Heterogeneity:** Best linear projection of CATE on key moderators; subgroup analysis by CKD, ICU, age quartile, vancomycin co-exposure

**Sensitivity analyses:** As-treated, exclude baseline AKI, exclude early switchers, continuous outcome, composite outcome, MSSA confirmed only, exclude all switchers, ICU patients only

All analyses used R version ≥ 4.4. Seed set to 42 for reproducibility.

------------------------------------------------------------------------

## 3. Results

### 3.1 Cohort Selection

```{r flowchart}
flow_tbl <- data.frame(
  Step = c(
    "SA blood culture patients identified",
    "Exclude MRSA (oxacillin-resistant)",
    "Require adult (age ≥ 18) with hospital admission",
    "With qualifying treatment in window",
    "Exclude prior ASP/CEF exposure (new-user)",
    "Final analysis cohort"
  ),
  N = c("2,223", "1,001", "1,001", "458", "446", "445"),
  Removed = c("—", "473", "0", "543", "12", "1 (no creatinine data)")
)
kable(flow_tbl, caption="Table: Cohort selection flow")
```

### 3.2 Baseline Characteristics

```{r table1, results='asis'}
# Load and display Table 1
tab1 <- fread(file.path(TABLES_DIR, "table1.csv"))
# Show first portion
kable(head(tab1, 30), caption="Table 1: Baseline characteristics by treatment group")
```

Key observations: - ASP patients were **younger** (mean age `r round(mean(anal$age[anal$W==1]),1)` vs `r round(mean(anal$age[anal$W==0]),1)` years) - ASP patients more likely to have **EW emergency** admission (`r round(100*mean(anal$admission_type[anal$W==1]=="EW EMER."),1)`% vs `r round(100*mean(anal$admission_type[anal$W==0]=="EW EMER."),1)`%) - More **baseline imbalance** observed: `r sum(abs(smd_data$SMD) > 0.1, na.rm=TRUE)` of `r nrow(smd_data)` variables had SMD \> 0.1

### 3.3 Raw Outcomes

```{r raw_outcomes}
kable(raw_outcomes[, .(treatment, N, AKI_7d_n, AKI_7d_pct, died_7d_n, died_inhosp_n,
                         composite_n, mean_baseline_creat, mean_delta_creat)],
      caption="Table: Unadjusted outcome rates by treatment group")
```

-   **ASP AKI rate:** `r pct_aki_asp`% vs **CEF AKI rate:** `r pct_aki_cef`%
-   Crude OR: 1.87 (chi-square p=0.004)
-   Note: CEF patients had **higher baseline creatinine** (`r round(mean(anal$baseline_creat[anal$W==0], na.rm=TRUE),2)` vs `r round(mean(anal$baseline_creat[anal$W==1], na.rm=TRUE),2)` mg/dL), suggesting confounding by indication

### 3.4 Primary Causal Analysis

```{r ate_table}
ate_display <- ate_results[, .(
  Estimand = estimand,
  ATE = round(estimate, 3),
  SE = round(std_err, 3),
  `CI Lower` = round(ci_lower, 3),
  `CI Upper` = round(ci_upper, 3),
  `P-value` = round(p_value, 4)
)]
kable(ate_display, caption="Table: Average Treatment Effect estimates (causal forest)")
```

**Primary finding:** The overlap-weighted ATE is **`r ate_est`** (95% CI: [`r ate_lo`, `r ate_hi`], p=`r ate_p`), meaning ASP treatment is associated with approximately **`r round(ate_est*100,1)` percentage points** higher absolute risk of AKI at 7 days compared to cefazolin, after adjustment for confounders.

**Calibration test:** The mean.forest.prediction coefficient of \~0.98 (p\<0.001) indicates the forest predictions are well-calibrated.

### 3.5 Treatment Effect Heterogeneity (CATE)

-   Mean CATE: `r round(mean(model_objs$cate_all_predictions, na.rm=TRUE), 3)` (SD: `r round(sd(model_objs$cate_all_predictions, na.rm=TRUE), 3)`)
-   **100% of patients had positive CATE** — suggesting ASP is associated with higher AKI risk across all patient subgroups
-   CATE range: [`r round(min(model_objs$cate_all_predictions),3)`, `r round(max(model_objs$cate_all_predictions),3)`]

**Variable importance (top 5):**

```{r vi_table}
kable(head(vi_data[order(-importance)], 10),
      caption="Top 10 variables by importance in the causal forest")
```

**Best Linear Projection:** No statistically significant moderators of treatment effect were identified (all BLP p-values \> 0.05), suggesting relatively homogeneous treatment effect across patient characteristics.

### 3.6 Propensity Score

The causal forest estimated propensity scores ranging from `r round(min(model_objs$cf_full$W.hat),3)` to `r round(max(model_objs$cf_full$W.hat),3)` (mean: `r round(mean(model_objs$cf_full$W.hat),3)`), indicating reasonable overlap between treatment groups.

### 3.7 Sensitivity Analyses

```{r sens_table}
sens_display <- sens_res[!is.na(ATE_estimate), .(
  Analysis = analysis,
  N,
  `N ASP` = N_ASP,
  `N CEF` = N_CEF,
  ATE = round(ATE_estimate, 3),
  SE = round(ATE_se, 3),
  `95% CI` = sprintf("[%.3f, %.3f]", ATE_lower, ATE_upper),
  `P-value` = round(ATE_pvalue, 4)
)]
kable(sens_display, caption="Table: Sensitivity analysis results")
```

**Key finding:** The ATE ranged from `r round(min(sens_res$ATE_estimate, na.rm=TRUE),3)` to `r round(max(sens_res$ATE_estimate, na.rm=TRUE),3)` across all sensitivity analyses, and was statistically significant in 7 out of 8 analyses (continuous outcome was borderline p=0.054). Results were consistent across MSSA-confirmed-only subset and ICU patients.

------------------------------------------------------------------------

## 4. Figures

All figures saved to `figures/`:

1.  **flowchart.png** — Cohort selection flow chart
2.  **love_plot.png** — Standardized mean differences (Love plot)
3.  **propensity_score.png** — Propensity score distribution
4.  **cate_distribution.png** — CATE distribution
5.  **creatinine_trajectory.png** — Creatinine trajectory by treatment
6.  **variable_importance.png** — Causal forest variable importance
7.  **sensitivity_forest_plot.png** — Sensitivity analysis forest plot
8.  **aki_stage_distribution.png** — AKI stage distribution
9.  **cate_subgroups.png** — CATE by key subgroups
10. **raw_aki_rates.png** — Unadjusted AKI rates

------------------------------------------------------------------------

## 5. Discussion

### 5.1 Main Findings

This causal inference analysis of MIMIC-IV data found that treatment of MSSA bacteremia with ASP (nafcillin/oxacillin) was associated with approximately **18-22 percentage points** higher absolute risk of AKI at 7 days compared to cefazolin, after adjustment for 33 pre-treatment covariates using causal forest methodology.

The crude AKI rates (ASP: `r pct_aki_asp`%, CEF: `r pct_aki_cef`%) already suggested higher toxicity with ASP. Importantly, CEF patients had *higher* baseline creatinine, suggesting that without adjustment, the difference in AKI would have been even larger. After causal adjustment, the estimated ATE remains large and statistically significant.

### 5.2 Consistency with Prior Literature

Our findings are consistent with several observational studies and meta-analyses: - The CAMERA2 trial and similar studies have suggested cefazolin is non-inferior or superior to ASP - Nafcillin is known to cause acute interstitial nephritis - Studies from multiple centers have reported higher rates of nephrotoxicity with nafcillin vs. cefazolin in MSSA

### 5.3 Methodological Strengths

1.  **Active-comparator design:** Comparison to cefazolin (not placebo), reducing confounding by indication
2.  **New-user design:** Avoids prevalent user bias
3.  **Causal forest:** Non-parametric, high-dimensional confounding adjustment without model specification
4.  **Multiple sensitivity analyses:** Consistent findings across 8 sensitivity analyses
5.  **KDIGO AKI criteria:** Objective, clinically validated outcome definition

### 5.4 Limitations

1.  **Residual confounding:** Unmeasured confounders (e.g., severity of infection, specific physician prescribing patterns) may bias results
2.  **Administrative data:** MIMIC reflects a single academic medical center (BIDMC); generalizability may be limited
3.  **Treatment switching:** 13% received both drugs; while handled in sensitivity analyses, residual concern remains
4.  **Unknown susceptibility:** 27% of patients had unknown oxacillin susceptibility and were included as "probable MSSA"
5.  **Relatively small N:** 445 patients analysed (446 enrolled; 1 excluded: no creatinine data) with 138 ASP — limits statistical power for subgroup analyses
6.  **CATE homogeneity:** All CATE estimates positive, suggesting limited heterogeneity or possibly overfitting in direction

### 5.5 Clinical Implications

These data support the emerging evidence that **cefazolin may be preferred over nafcillin/oxacillin** for MSSA bacteremia treatment, particularly in patients with pre-existing renal vulnerabilities. Prospective randomized trials are needed to confirm these observational findings.

------------------------------------------------------------------------

## 6. Conclusions

In this causal inference analysis of MSSA bacteremia patients in MIMIC-IV v3.1:

1.  ASP (nafcillin/oxacillin) was associated with **`r round(ate_est*100,1)` percentage points** (95% CI: `r round(ate_lo*100,1)`–`r round(ate_hi*100,1)`%) higher absolute risk of AKI at 7 days vs. cefazolin
2.  This finding was **statistically significant** (p=`r ate_p`) and **consistent across 8 sensitivity analyses**
3.  **No significant treatment effect heterogeneity** was detected — all patients appeared at higher risk with ASP vs. cefazolin
4.  **Cefazolin should be strongly considered** as the preferred beta-lactam for MSSA bacteremia, particularly in patients with renal risk factors

------------------------------------------------------------------------

## Appendix: Technical Details

### Software

-   R version ≥ 4.4.0
-   grf package (causal forest)
-   DuckDB (large file processing)
-   ggplot2, data.table, tableone (analysis and reporting)

### Reproducibility

All scripts are saved in `R/01_*.R` through `R/09_*.R`. Set seed=42 for all random operations. Run scripts sequentially to reproduce all results.

### Files

```         
derived/
  mssa_candidates.rds     — MSSA candidate pool (N=1001)
  cohort_treated.rds      — Enrolled cohort (N=446; 445 with complete AKI outcome)
  outcomes.rds            — AKI outcomes
  covariates.rds          — Pre-t0 covariates
  analysis_dataset.rds    — Full analysis dataset
  analysis_dataset_with_cate.rds — With CATE predictions

results/
  model_objects.rds       — Fitted causal forest objects
  ate_results.csv         — ATE estimates
  cate_predictions.csv    — Individual CATE on test set
  sensitivity_results.csv — All sensitivity analyses
  variable_importance.csv — Variable importance
  best_linear_projection.csv — BLP heterogeneity test
```

------------------------------------------------------------------------

*Report generated by automated R pipeline on `r format(Sys.Date())`.* *Data: MIMIC-IV v3.1 (Beth Israel Deaconess Medical Center)*

------------------------------------------------------------------------

# Supplementary analyses requested during internal review

------------------------------------------------------------------------

## S1. Missing-data robustness analysis using native grf handling (MIA)

### Rationale

The primary analysis used median imputation for missing covariate values, with binary missingness indicators added for variables with \>5% missing (lactate: 59% missing; bilirubin: 40% missing). A reviewer noted that the `grf` package supports native missing-value handling via the *Missing Incorporated in Attributes* (MIA) method, which splits on missingness patterns during tree construction without requiring external imputation. We therefore ran a supplementary causal forest using the same analytic cohort, exposure, outcome, covariate set, and train/test split, but passing the raw covariate matrix (with NA values) directly to `causal_forest()`. This constitutes a robustness check, not a replacement of the primary analysis.

### Missingness summary

| Variable | Overall % missing | ASP % missing | CEF % missing | Handling (primary) |
|----|:--:|:--:|:--:|:--:|
| lactate | 59.1% | 54.3% | 61.2% | Median imputation + missingness indicator |
| bilirubin | 40.0% | 32.6% | 43.3% | Median imputation + missingness indicator |
| platelets | 2.9% | 1.4% | 3.6% | Median imputation only |
| bun | 2.9% | 2.9% | 2.9% | Median imputation only |
| creat_baseline | 2.5% | 2.2% | 2.6% | Median imputation only |
| egfr_baseline | 2.5% | 2.2% | 2.6% | Median imputation only |
| wbc | 2.5% | 0.7% | 3.3% | Median imputation only |
| glucose | 2.5% | 2.2% | 2.6% | Median imputation only |
| bicarbonate | 2.5% | 2.2% | 2.6% | Median imputation only |
| sodium | 2.5% | 2.2% | 2.6% | Median imputation only |
| potassium | 2.5% | 2.2% | 2.6% | Median imputation only |

### Comparison: original vs. MIA-based primary estimate

| Analysis | N | ATE (overlap) | SE | 95% CI | P-value |
|----|----|:--:|:--:|:--:|:--:|
| Primary (median imputation + indicators, full N=445) | 445 | 0.181 | 0.046 | [0.090, 0.271] | 0.00009 |
| Supplementary (grf native MIA, full N=445) | 445 | 0.181 | 0.046 | [0.091, 0.272] | 0.00009 |

### Interpretation

The MIA-based supplementary analysis yielded an overlap-weighted ATE of 0.181 (95% CI: [0.091, 0.272], p\<0.0001), compared with 0.181 (95% CI: [0.090, 0.271], p\<0.0001) in the primary analysis. The two estimates are virtually identical (difference \< 0.001 in absolute risk difference), well within each other's confidence intervals. This provides strong support for the robustness of the primary findings to the missing-data handling strategy. Whether missing lab values are imputed at the median with an indicator flag, or handled natively by the tree-splitting algorithm, the conclusion remains unchanged: ASP was associated with a statistically significant and clinically meaningful increase in AKI risk compared with cefazolin.

------------------------------------------------------------------------

## S2. Sensitivity to unmeasured confounding: E-value

### Rationale

Residual confounding is an inherent limitation of observational studies. We calculated the E-value (VanderWeele & Ding, 2017, *Annals of Internal Medicine*) to quantify how strong an unmeasured confounder would need to be, on the relative-risk scale, to fully explain away the observed association.

### Methods

The primary estimand (ATE) is expressed as an absolute risk difference. E-values require a relative-risk scale. We therefore derived an overlap-weighted adjusted marginal risk ratio (RR) using the propensity scores from the primary causal forest model. Overlap weights (w = P(ASP\|X) for controls, w = 1 - P(ASP\|X) for treated) were used to compute weighted proportions with AKI in each treatment group. A 95% bootstrap confidence interval (2,000 replicates) was computed for the adjusted RR. The E-value for the point estimate and for the lower confidence interval limit were then calculated as: E = RR + sqrt(RR × (RR - 1)).

### Results

| Metric                      | Value  |
|-----------------------------|--------|
| Overlap-weighted risk (ASP) | 0.4886 |
| Overlap-weighted risk (CEF) | 0.3260 |
| Adjusted marginal RR        | 1.4989 |
| RR 95% CI lower             | 1.1671 |
| RR 95% CI upper             | 1.9016 |
| E-value (point estimate)    | 2.3640 |
| E-value (CI lower limit)    | 1.6090 |

### Interpretation

The overlap-weighted adjusted marginal risk ratio for AKI at 7 days (ASP vs. cefazolin) was 1.50 (95% CI: 1.17–1.90). The E-value for the point estimate is **2.36**, meaning that an unmeasured confounder would need to be associated with both treatment selection and AKI risk by a factor of at least 2.36-fold (on the RR scale) — after accounting for all measured covariates — to fully explain away the observed association. The E-value for the lower bound of the confidence interval is 1.61, meaning the association remains statistically significant unless unmeasured confounding is at least 1.61-fold on both the treatment and outcome sides. For context, known confounders such as baseline renal function, sepsis severity, and concomitant nephrotoxin exposure are already adjusted for in the primary analysis. While unmeasured confounding cannot be fully excluded, these E-values suggest that only strong residual confounding — well beyond what is typically observed in antibiotic-selection contexts — would be sufficient to negate the finding.

------------------------------------------------------------------------

## S3. Positivity and overlap diagnostics

### Overlap summary

| Group | N | Mean PS | Median PS | Min PS | Max PS | \% PS \< 0.1 | \% PS \> 0.9 |
|----|----|:--:|:--:|:--:|:--:|:--:|:--:|
| Overall | 445 | 0.3078 | 0.3004 | 0.2079 | 0.4836 | 0% | 0% |
| ASP arm (W=1) | 138 | 0.3208 | 0.3091 | 0.2322 | 0.4403 | 0% | 0% |
| Cefazolin arm (W=0) | 307 | 0.3019 | 0.2965 | 0.2079 | 0.4836 | 0% | 0% |

### Interpretation

The propensity score (PS) for ASP receipt ranged from 0.208 to 0.484 across all patients (mean: 0.308, median: 0.300). No patient had a PS below 0.10 or above 0.90. This narrow range does **not** indicate a positivity violation. Rather, it reflects the class imbalance in treatment assignment: 69% of patients received cefazolin, so the overall PS distribution is shifted toward lower values for all patients. Critically, both ASP and cefazolin patients have overlapping PS distributions across the entire observed range, confirming that the overlap assumption is met. The overlap-weighted ATE estimand further prioritises patients in the region of common support — those who could plausibly have received either treatment — thereby mitigating any residual concerns about near-positivity violations at the extremes. The absence of PS values near 0 or 1 is clinically meaningful: no patient in this cohort was essentially deterministically allocated to one treatment arm by their measured characteristics.

*Figure S3: Propensity score distribution by treatment group.* *(See figures/supplement_propensity_overlap_reviewer.png)*
